---
title: "Get started with `merra2ools`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{merra2ools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE, 
  message = FALSE,
  comment = "#>")
knitr::opts_chunk$set(fig.width = 7, fig.height = 6, fig.align = "center") 
library(knitr)

```

## Installation
**merra2ools** is package for [R](https://www.r-project.org/) and a dataset. It is assumed that R version >=4.0 is pre-installed. ([`Rtools`](https://cran.r-project.org/bin/windows/Rtools/) is also required on Windows to build this and several other packages from source or GitHub). It is also recommended (though it is not required) to use [RStudio](https://www.rstudio.com/products/rstudio/download/) or other IDE for R. `merra2ools` depends on several R-packages which (if not yet available on your system) will be installed along with the `merra2ools` installation (see the package [DESCRIPTION](https://github.com/energyRt/merra2ools/blob/master/DESCRIPTION) for details). The package also requires its dataset to be downloaded separately from the installation of the package (see below). Though `merra2ools` package can operate without the dataset (for estimation capacity factors, solar geometry, etc.), the supplied data should be formatted in the same way as it is expected by the package functions.   
\    
**merra2sample** is 12-days example of the the `merra2ools` subset (41 years), 21st-day of each month of 2010. It can installed directly from GitHub, and used for quick checks of the data and its format; it is also used to build vignettes (articles) of `merra2ools` package. Since the example dataset has considerable size for R-packages (~0.5Gb) and mostly repeats the main dataset, it is saved as a separate package.  

```{r, results=FALSE}
pkg <- function() rownames(installed.packages()) # returns names of installed packages
# Installation of `merra2ools` package
if (!("remotes" %in% pkg())) install.packages("remotes")
if (!("merra2ools" %in% pkg())) remotes::install_github("energyRt/merra2ools", dependencies = TRUE)
if (!("merra2sample" %in% pkg())) remotes::install_github("energyRt/merra2sample")
# Packages used in the vignette
if (!("rnaturalearthhires" %in% pkg()))
  devtools::install_github("ropensci/rnaturalearthhires")
if (!("scales" %in% pkg())) install.packages("scales")
if (!("cowplot" %in% pkg())) install.packages("cowplot")
```

**`merra2ools` dataset** (~270Gb) can be downloaded from [https://doi.org/10.5061/dryad.v41ns1rtt](https://doi.org/10.5061/dryad.v41ns1rtt), unziped in a dedicated directory on an internal or external hard-drive, and configured as described below.  

## Setting up  
Loading packages used in the vignette.
```{r setup, eval=T, results=FALSE, warning=F, message=FALSE}
library(tidyverse)
library(merra2ools)
library(sp)
library(sf)
library(cowplot)
```

Checking if the dataset is connected.
```{r}
check_merra2()
```

Connecting the dataset (if not yet connected)
```{r, eval=FALSE}
check_merra2("RAPLACE WITH PATH TO THE DOWNLOADED DATA") # check if the data in the directory
set_merra2_options(merra2.dir = "RAPLACE WITH PATH TO THE DOWNLOADED DATA") # safe the path
get_merra2_dir()  # check if the path is saved
check_merra2(detailed = T)
```

## Accessing the data  
The database is organized in monthly files in `fst` format. There are two functions to read a whole file (`read_merra_file()`) or read a subset for specified locations and time period (`get_merra2_subset()`). Both functions provide an option to read the dataset in the scaled format, in reported units (default) or "raw" format (as integers - the way it is stored).  

```{r, eval=T}
x <- read_merra_file("202012", as_integers = TRUE)
```

```{r, echo=FALSE}
if (!("kableExtra" %in% pkg())) install.packages("kableExtra")
library(kableExtra)
kable(x[1:5,], caption = "Subset of raw data (as integers)") %>%
  kable_styling(font_size = 8, full_width = T) %>%
  column_spec(1, width_min = "11em", background = "lightgrey") %>%
  column_spec(2, background = "lightgrey")
```

```{r, eval=FALSE}
x <- get_merra2_subset(from = "20201130 01", to = "20201201 23")
```

```{r, echo=FALSE}
# if (!("kableExtra" %in% pkg())) install.packages("kableExtra")
# library(kableExtra)
kable(x[1:5,], caption = "merra2ools subset") %>%
  kable_styling(font_size = 8, full_width = T) %>%
  column_spec(1, width_min = "11em", background = "lightgrey") %>%
  column_spec(2, background = "lightgrey")
```

## Locations identifiers  
Every data-point in the used MERRA-2 collections (see the dataset description) is associated with coordinates and time. The original MERRA-2 files have index-variables `V1`, `V2`, and `V3` to identify longitude, latitude, and time (hour) dimensions respectively. For convenience, a location identifier `locid` has been generated as a Kronecker product of `V1` and `V2`. The `locid` identifier is used as the key variable of MERRA-2 locations, instead of `V1` and `V2`. The time identifier (`V3` - hour) is combined with the year, month, and day in `UTC` key variable. The total number of location points in MERRA-2 is 207,936 (576 X 361). The identifier is saved in the `merra2ools` package `\data` directory and can be called with `data` function (`data("locid")`) or directly `locid`.
```{r, eval=F}
# Ways to call `locid` data.frame
data("locid")
merra2ools::locid
locid
```
\   
```{r, echo=FALSE}
kable(rbind(locid[1:3,], locid[207934:207936,]),
      caption = "Location identifiers", label = "Table") %>%
  kable_styling(font_size = 8, position = "center")
```

<details><summary>code</summary>
```{r, eval=FALSE}
lon <- unique(locid$lon)
head(lon, 10)
length(lon)

lat <- unique(locid$lat)
head(lat, 10)
length(lat)

lo <- unique(c(seq(-180, max(lon), by = 30), max(lon))) %>% sort()
la <- seq(-90, 90, by = 15)

locid_sample <- filter(locid, lon %in% lo, lat %in% la)
world_map <- rnaturalearth::ne_countries(scale = "small", returnclass = "sp")
world_map <- fortify(world_map) %>% as_tibble()

fig.locid <- ggplot() + 
  geom_map(data = world_map, map = world_map, 
           aes(x = long, y = lat, map_id = id), 
           fill = "wheat", alpha = .75, colour = NA, size = 0.25) +
  theme_bw() +
  geom_rect(aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
    data = data.frame(xmin = -180, xmax = 180, ymin = -90, ymax = 90),
    alpha = .5, fill = NA, colour = "grey85") +
  geom_point(aes(x = lon, y = lat), data = locid_sample, size = 1, colour = "red") + 
  geom_text(aes(x = lon, y = lat, label = locid), data = locid_sample, 
             position = position_nudge(y = 4),
             alpha = 0.75,
             size = unit(3, "lines")) +
  scale_x_continuous(breaks = ceiling(lo), minor_breaks = ceiling(lo)) +
  scale_y_continuous(breaks = round(la), minor_breaks = ceiling(la)) + 
  labs(x = "lon") +
  # labs(x = "lon", title = "Location ID (`locid`) layout of MERRA-2 subset") +
  theme(plot.title = element_text(hjust = 0.5)) + 
  coord_fixed(1.)

ggsave("images/locid_map.png", fig.locid, width = 9, height = 5.)

```
</details>

```{r fig_MERRA_locid, fig.cap="locid_map", echo=FALSE}
fig_file <- "../images/locid_map.png"
if (file.exists(fig_file)) include_graphics(fig_file)
```

## Matching `locid` with a map  
Location identifiers in `merra2ools` dataset can be seen as spatial points or centers (centroids) of a spatial grid or spatial polygons. Subsetting `locid` for a particular geographical region can be done using a "map" of the region in `SpatialPolygonsDataFrame` format (`spdf`). Function `get_locid` offers two alternative criteria of selecting `locid`s is implemented in `get_locid` function:  
  * `locid` as a spatial points *overlay* the map's spatial polygons (method = "points");   
  * `locid` as a spatial polygon *intersect* with the map's spatial polygons (method = "intersect").  

\   
Three examples below compare subsetting with the two methods for three different regions/countries.   


### Example 1. Florida, USA   
```{r, results=F, eval=FALSE}
# US-map
usa_sp <- rnaturalearth::ne_states(iso_a2 = "us")

# Subset Florida map (keep two formats - `spdf` and `sf`)
florida_sp <- usa_sp[usa_sp$name_en == "Florida",]
florida_sf <- st_as_sf(florida_sp[, "name"])

# location IDs, two methods
locid_fl_p <- get_locid(florida_sp, method = "points")
locid_fl_i <- get_locid(florida_sp, method = "intersect")

# MERRA-2 grid for the selected `locid's`
locid_fl_grid <- getGrid("poly", locid = locid_fl_i)
# locid_fl_sp <- get_locid(florida_sp, return_sp = T)

# Plots
png("images/example1_florida_locids.png", width = 5, height = 5, units = "in", res = 200)
sp::plot(florida_sp, axes = T, col = "wheat")
sp::plot(locid_fl_grid, add = T, border = "darkgrey")
points(locid$lon[locid_fl_p], locid$lat[locid_fl_p], col = "red", pch = 16)
points(locid$lon[locid_fl_i], locid$lat[locid_fl_i], col = "red", pch = 1)
dev.off()
```

```{r fig_florida_locid, fig.cap="MERRA-2 locations for Florida", echo=FALSE}
fig_file <- "../images/example1_florida_locids.png"
if (file.exists(fig_file)) include_graphics(fig_file)
```


### Example 2. Iceland

<details><summary>code</summary>
```{r, results=F, eval=FALSE}
# Map
# iceland_sp <- rnaturalearth::ne_states(country = "iceland")
iceland_sp <- rnaturalearth::ne_countries(country = "iceland", scale = "large")
iceland_sf <- st_as_sf(iceland_sp[, "name"])

str(iceland_sp@data)
# location IDs, two methods
locid_isl_p <- get_locid(iceland_sp, method = "points")
locid_isl_i <- get_locid(iceland_sp, method = "intersect")

# MERRA-2 grid for the selected `locid's`
locid_isl_grid <- getGrid("poly", locid = locid_isl_i)

# Plots
png("images/example2_iceland_locids.png", width = 5, height = 5, units = "in", res = 200)
sp::plot(iceland_sp, axes = T, col = "wheat")
sp::plot(locid_isl_grid, add = T, border = "darkgrey")
points(locid$lon[locid_isl_p], locid$lat[locid_isl_p], col = "red", pch = 16)
points(locid$lon[locid_isl_i], locid$lat[locid_isl_i], col = "red", pch = 1)
dev.off()
```
</details>

```{r fig_iceland_locid, fig.cap="MERRA-2 locations for Iceland", echo=FALSE}
fig_file <- "../images/example2_iceland_locids.png"
if (file.exists(fig_file)) include_graphics(fig_file)
```


### Example 3. Kenya

<details><summary>code</summary>
```{r, results=F, eval=FALSE}
# Map
# kenya_sp <- rnaturalearth::ne_states(country = "kenya")
kenya_sp <- rnaturalearth::ne_countries(country = "kenya", scale = "large")
kenya_sf <- st_as_sf(kenya_sp[, "name"])

# location IDs, two methods
locid_ken_p <- get_locid(kenya_sp, method = "points")
locid_ken_i <- get_locid(kenya_sp, method = "intersect")

# MERRA-2 grid for the selected `locid's`
locid_ken_grid <- getGrid("poly", locid = locid_ken_i)

# Plots
png("images/example3_kenya_locids.png", width = 5, height = 5, units = "in", res = 200)
sp::plot(kenya_sp, axes = T, col = "wheat", main = "Kenya")
sp::plot(locid_ken_grid, add = T, border = "darkgrey")
points(locid$lon[locid_ken_p], locid$lat[locid_ken_p], col = "red", pch = 16, cex = .75)
points(locid$lon[locid_ken_i], locid$lat[locid_ken_i], col = "red", pch = 1, cex = .75)
dev.off()
```
</details>


```{r fig_kenya_locid, fig.cap="MERRA-2 locations for Kenya", echo=FALSE}
fig_file <- "../images/example3_kenya_locids.png"
if (file.exists(fig_file)) include_graphics(fig_file)
```

## Get MERRA-2 subset
```{r, eval=FALSE}
merra_fl <- get_merra2_subset(locid = locid_fl_i, 
                              from = "20190101 00", to = "20191231 23", 
                              tz = "US/Eastern")

merra_isl <- get_merra2_subset(locid = locid_isl_i, 
                               from = "20190101 00", to = "20191231 23", 
                               tz = "Atlantic/Reykjavik")

merra_ken <- get_merra2_subset(locid = locid_ken_i, 
                               from = "20190101 00", to = "20191231 23", 
                               tz = "Africa/Nairobi")
```

## Wind power 
Estimation of wind power capacity factors (CF) for the first example - *Florida*.  
```{r, eval=FALSE}
# Estimate capacity factors
merra_fl <- fWindCF(merra_fl, height = 150)

# Annual averages
win_fl_y <- merra_fl %>% # annual averages
  group_by(locid) %>%
  summarise(win150af = mean(win150af, na.rm = T)) %>%
  add_coord()
```

Do the same for other examples - Iceland and Kenya.  
<details><summary>code</summary>
**Iceland**
```{r, eval=FALSE}
# Estimate capacity factors
merra_isl <- fWindCF(merra_isl, height = 150)

# Annual averages
win_isl_y <- merra_isl %>% # annual averages
  group_by(locid) %>%
  summarise(win150af = mean(win150af, na.rm = T)) %>%
  add_coord()
```

**Kenya**
```{r, eval=FALSE}
# Estimate capacity factors
merra_ken <- fWindCF(merra_ken, height = 150)

# Annual averages
win_ken_y <- merra_ken %>% 
  group_by(locid) %>%
  summarise(win150af = mean(win150af, na.rm = T)) %>%
  add_coord()
```
</details>

## Solar power
Estimation of Plane of Array Irradiance (POA) for fixed tilted (fl) array-systems for *Florida*  
```{r, eval=FALSE}
# Estimate POA
merra_fl <- merra_fl %>% fPOA(array.type = "fl")

# Annual averages
poa_fl_y <- merra_fl %>%
  group_by(locid) %>%
  summarise(POA.fl = sum(POA.fl, na.rm = T)/365/1e3) %>%
  add_coord() %>%
  as_tibble()
```

Repeat for other examples.  
<details><summary>code</summary>
**Iceland**
```{r, eval=FALSE}
# Estimate POA
merra_isl <- merra_isl %>% fPOA(array.type = "fl")

# Annual averages
poa_isl_y <- merra_isl %>%
  group_by(locid) %>%
  summarise(POA.fl = sum(POA.fl, na.rm = T)/365/1e3) %>%
  add_coord() %>%
  as_tibble()
```

**Kenya**
```{r, eval=FALSE}
# Estimate POA
merra_ken <- merra_ken %>% fPOA(array.type = "fl")

# Annual averages
poa_ken_y <- merra_ken %>%
  group_by(locid) %>%
  summarise(POA.fl = sum(POA.fl, na.rm = T)/365/1e3) %>%
  add_coord() %>%
  as_tibble()
```

**Comparative figure**   
To use the same scale, 
```{r, eval=FALSE}
wnd_range <- range(win_fl_y$win150af, win_isl_y$win150af, win_ken_y$win150af)
wnd_breaks <- scales::breaks_pretty(5)(wnd_range)
wnd_range <- range(wnd_breaks)

poa_range <- range(poa_fl_y$POA.fl, poa_isl_y$POA.fl, poa_ken_y$POA.fl)
poa_breaks <- scales::breaks_pretty(5)(poa_range)
poa_range <- range(poa_breaks)

```

```{r, eval=FALSE}
# Plot capacity factor variable on map
cf_plot <- function(
  data, gis_sf, var_name, brakes, 
  labels = brakes, limits = range(brakes), legend_name = var_name,
  viridis_palette = "viridis", border_colour = alpha("white", .5),
  legend_position = "none") {
  ggplot() +
    geom_tile(aes_string("lon", "lat", fill = var_name), data = data) +
    scale_fill_viridis_c(
      breaks = brakes, labels = labels, limits = limits, name = legend_name,
      option = viridis_palette) +
    geom_sf(data = gis_sf, fill = NA, colour = border_colour) +
    theme_bw() + theme(legend.position = legend_position) +
    theme(plot.margin = margin(0.5, 0.0, 0.0, 0.0, "cm")) +
    labs(x = "", y = "")
}

# Combine plots into one
plot_grid(
  cf_plot(win_fl_y, florida_sf, "win150af", wnd_breaks),
  cf_plot(win_isl_y, iceland_sf, "win150af", wnd_breaks),
  cf_plot(win_ken_y, kenya_sf, "win150af", wnd_breaks,
          legend_position = "right", legend_name = "Wind\nCF"),
  # NULL, NULL, NULL,  
  cf_plot(poa_fl_y, florida_sf, "POA.fl", poa_breaks, viridis_palette = "plasma"),
  cf_plot(poa_isl_y, iceland_sf, "POA.fl", poa_breaks, viridis_palette = "plasma"),
  cf_plot(poa_ken_y, kenya_sf, "POA.fl", poa_breaks, viridis_palette = "plasma",
          legend_position = "right", legend_name = "POA\nkW/day"),
  # rel_heights = c(1, -.3, 1),
  ncol = 3, rel_widths = c(1, 1.25, 1.23),
  labels = c("Florida", "Iceland", "Kenya"), hjust = c(-1.6, -1.9, -1.6)
)
ggsave2("images/cf_group.png", width = 8, height = 5, dpi = 200)

```
</details>


```{r fig_cf_group, fig.cap="Wind power capacity factors (CF) and Solar irradiance on Plane of Array (POA) by regions", echo=FALSE}
fig_file <- "../images/cf_group.png"
if (file.exists(fig_file)) include_graphics(fig_file)
```

## Precipitation by months   

```{r, eval=FALSE}
prec_ken_m <- merra_ken %>%
  mutate(
    local_time = lubridate::with_tz(UTC, "Africa/Nairobi"),
    year = year(local_time),
    month = month(local_time),
    month_name = factor(month.name[month], levels = month.name, ordered = TRUE)) %>%
  group_by(locid, year, month, month_name) %>%
  summarise(
    PRECTOTCORR = sum(PRECTOTCORR, na.rm = T),
    T10M = mean(T10M)) %>%
  add_coord() %>%
  as_tibble()
```

<details><summary>code</summary>

```{r, eval=FALSE}
pre_range <- range(prec_ken_m$PRECTOTCORR)
pre_breaks <- scales::breaks_pretty(5)(pre_range)
pre_range <- range(pre_breaks)

cf_plot(prec_ken_m, kenya_sf, "PRECTOTCORR", brakes = pre_breaks, legend_position = "right") + 
  facet_wrap(.~month)

ggplot(prec_ken_m) +
  geom_tile(aes(lon, lat, fill = PRECTOTCORR)) + 
  scale_fill_distiller(palette = "YlGnBu", direction = 1, name = "mm /\nmonth") +
  facet_wrap(.~month_name) + 
  geom_sf(data = kenya_sf, fill = NA, colour = alpha("black", .5)) +
  theme_bw() + labs(x = "", y = "")
ggsave("images/precipitation_m_kenya.png", scale = 1.5)

ggplot(prec_ken_m) +
  geom_tile(aes(lon, lat, fill = T10M)) + 
  scale_fill_distiller(palette = "Spectral", direction = -1, name = "\u00B0C") +
  facet_wrap(.~month_name) + 
  geom_sf(data = kenya_sf, fill = NA, colour = alpha("black", .5)) +
  theme_bw()
ggsave("images/temperature_m_kenya.png", scale = 1.5)
```
</details>

```{r fig_prec_ken, fig.cap="Precipitation by month in Kenya (2019), millimeters / hour (kg/m^2/hour)", echo=FALSE}
fig_file <- "../images/precipitation_m_kenya.png"
if (file.exists(fig_file)) include_graphics(fig_file)
```

## Air temperature by months   

```{r fig_temp_ken, fig.cap="Average monthly temperature in Kenya (2019), \u00B0C)", echo=FALSE}
fig_file <- "../images/temperature_m_kenya.png"
if (file.exists(fig_file)) include_graphics(fig_file)
```

