---
title: "Solar Power"
output: rmarkdown::html_vignette
header-includes:
   - \usepackage{amsmath}
vignette: >
  %\VignetteIndexEntry{Solar Power}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

# Code execution parameters ####
# rerun <- TRUE # if TRUE, (re)runs the script and (re)saves all figures and gifs
rerun <- FALSE # if FALSE, knits the document using saved figures

```

```{r setup, eval = rerun}
library(merra2ools)
library(tidyverse)
library(cowplot)

# Example dataset
# merra <- merra2_sample(month = 1:12, add.coord = T)
# Note: month = 1:12 requires ~64Gb of RAM
#       one month (such as month = 3 or month = 1) requires ~10Gb or RAM
merra <- merra2_sample(month = c(3, 6, 9, 12), add.coord = T)
merra

```

## Plane of Array (POA) irradiance model
### All steps at once
```{r, eval = rerun}
x <- fPOA(merra)
# x <- fPOA(merra, keep.all = T) # keeps all interim variables

```


<details><summary>Figure code</summary>
```{r, eval = rerun}
fig.POA <- function(x, name, 
                    legend.position = NULL, 
                    timestamp.position = NULL, ...) {
  plot_merra(x, name = name, limits = c(0, 1200), 
           # legend.position = c(0.88, 0.05), 
           legend.position = legend.position,
           legend.name = ("W/\u33A1"), 
           timestamp.position = timestamp.position,
           expand.x = c(0.002, 0.005), expand.y = c(0.002, 0.005)
           )
}
fig.POA(x, "SWGDN")

fig.poa.grid <- function(x, 
                         legend.position = c(0.81, 0.08), 
                         timestamp.position = c(125, 85), 
                         ...) {
  plot_grid(fig.POA(x, "SWGDN", ...), 
            fig.POA(x, "POA.fh", timestamp.position = timestamp.position, ...),
            fig.POA(x, "POA.fl", ...), 
            fig.POA(x, "POA.th", ...),
            fig.POA(x, "POA.tv", ...), 
            fig.POA(x, "POA.td", legend.position = legend.position, ...),
            # fig.dPS(x),
            labels = c("GHI", "POA.fh", "POA.fl", "POA.th", "POA.tv", "POA.td"),
            ncol = 2, nrow = 3, hjust = -0.1, vjust = 1.1, label_size = 12,
            label_colour = "black")
}
fig.poa.grid(x)

ii <- rep(TRUE, nrow(x))
# ii <- lubridate::month(x$UTC) <= 1; summary(ii)
gif_merra(x[ii,], FUN = "fig.poa.grid", dirname = "images", fps = 5,
          gif.width = 1.25 * 576, 
          gif.height = 1.25 * 360 * 1.21, 
          filename.prefix = "POA_map_")

```
</details>

```{r fig_POA_all, fig.cap="Plane of Array (POA) Irradiance estimates for alternative tracking systems", echo=FALSE}
gif_file <- "../images/POA_map_720x544_5fps.gif"
if (file.exists(gif_file)) include_graphics(gif_file)

```

### Step-by-step
#### Solar position

```{r, eval = rerun}
x <- solar_position(merra, keep.all = T)

```

<details><summary>Figure code</summary>

```{r, eval = rerun}

x$na <- 1L; x$na[!x$beam] <- NA # drop nights

summary(x$beam)
plot_merra(x, "beam", limits = c(0, 1), palette = "YlGnBu",
           legend.name = "")

summary(x$declination)
fig.declination <- function(x, limits = c(-.5, .5), 
                            legend.name = "angle\u00b0", 
                            ...) {
  plot_merra(x, "declination", limits = limits, 
             # palette = "YlGnBu",
             legend.name = legend.name, ...)
}
fig.declination(x)

summary(x$eq_time)
fig.eq_time <- function(x, limits = c(-15, 15), ...) {
  plot_merra(x, "eq_time", limits = limits, palette = "PRGn", 
             legend.name = "Munutes", direction = -1, ...)
}
fig.eq_time(x)

summary(x$solar_time)
fig.solar_time <- function(x, limits = c(-12, 36), 
           palette = "Dark2", legend.name = "Hours", 
           ...) {
  plot_merra(x, "solar_time", limits = limits, 
             palette = palette, legend.name = legend.name, ...)
}
fig.solar_time(x)
  
summary(x$hour_angle)
summary(rad2deg(x$hour_angle))
fig.hour_angle <- function(x, limits = c(-360, 360), scale = 180/pi, 
           palette = "Set1", legend.name = "angle\u00b0", 
           direction = 1, ...) {
  plot_merra(x, "hour_angle", limits = limits, scale = scale,
             palette = palette, legend.name = legend.name, direction = direction, ...)
}
fig.hour_angle(x)

summary(x$zenith)
fig.zenith <- function(x, ...) {
  plot_merra(x, "zenith", limits = c(0, 90), 
             direction = 1, 
             scale = x$na,
             palette = "Blues", 
             legend.name = "angle\u00b0", ...)
}
fig.zenith(x)

summary(x$azimuth)
fig.azimuth <- function(x, ...) {
  plot_merra(x, "azimuth", limits = c(0, 360), 
             legend.name = "angle\u00b0",
             palette = "Set1",
             # palette = "Blues", 
             scale = x$na, direction = -1, ...)
}
fig.azimuth(x)

fig.solar_postion <- function(x, 
                         legend.position = c(0.80, 0.08), 
                         timestamp.position = c(125, 85), 
                         ...) {
  plot_grid(fig.declination(x, timestamp.position = NULL, 
                            legend.position = legend.position, ...), 
            fig.eq_time(x, timestamp.position = timestamp.position, 
                        legend.position = legend.position, ...),
            fig.solar_time(x, timestamp.position = NULL, 
                           legend.position = legend.position, ...), 
            fig.hour_angle(x, timestamp.position = NULL, 
                           legend.position = legend.position, ...),
            fig.zenith(x, timestamp.position = NULL, 
                       legend.position = legend.position, ...), 
            fig.azimuth(x, legend.position = legend.position, 
                        timestamp.position = NULL, ...),
            # fig.dPS(x),
            labels = c("declination", "eq_time", "solar_time", 
                       "hour_angle", "zenith", "azimuth"),
            ncol = 2, nrow = 3, hjust = -0.1, vjust = 1.1, label_size = 12,
            label_colour = "black")
}
fig.solar_postion(x)

# ii <- lubridate::hour(x$UTC) <= 2; summary(ii)
gif_merra(x, FUN = "fig.solar_postion", dirname = "images", fps = 5,
          gif.width = 1.25 * 576, 
          gif.height = 1.25 * 360 * 1.21, 
          filename.prefix = "fig.solar_postion_")

```
</details>


```{r fig_sol_pos, fig.cap="Solar position variables", echo=FALSE}
gif_file <- "../images/fig.solar_postion_720x544_5fps.gif"
if (file.exists(gif_file)) include_graphics(gif_file)

```


#### Global Horizontal Irradiance decomposition

<details><summary>Figure code</summary>
```{r, eval = rerun}
x <- ghi_decomposition(x, keep.all = T)

summary(x$ext_irrad)
fig.ext_irrad <- function(x, ...) {
  plot_merra(x, "ext_irrad", limits = c(1300, 1450), direction = 1,
             palette = "Oranges", legend.name = "W/\u33A1", ...)
}
fig.ext_irrad(x)

summary(x$clearness_index)
fig.clearness_index <- function(x, ...) {
  plot_merra(x, "clearness_index", limits = c(0, 1), direction = -1, palette = "Blues")
}
fig.clearness_index(x)

summary(x$diffuse_fraction)
fig.diffuse_fraction <- function(x, ...) {
  plot_merra(x, "diffuse_fraction", limits = c(0, 1), palette = "Blues", direction = 1)
}
fig.diffuse_fraction(x)

summary(x$DNI)
fig.DNI <- function(x, ...) {
  plot_merra(x, "DNI", limits = c(0, 1200), ...)
}
fig.DNI(x)

summary(x$DHI)
fig.DHI <- function(x, ...) {
  plot_merra(x, "DHI", limits = c(0, 1200), ...)
}
fig.DHI(x)

fig.ghi_decomposition <- function(x, 
                         legend.position = c(0.81, 0.08), 
                         timestamp.position = c(125, 85), 
                         ...) {
  plot_grid(fig.POA(x, "SWGDN", timestamp.position = NULL,
                    legend.position = legend.position, 
                    limits = c(0, 1500), ...),
            fig.ext_irrad(x, timestamp.position = timestamp.position, 
                          legend.position = legend.position, ...), 
            fig.clearness_index(
              x, timestamp.position = timestamp.position, 
              legend.position = legend.position, ...),
            fig.diffuse_fraction(x, timestamp.position = NULL, 
                           legend.position = legend.position, ...), 
            fig.DNI(x, timestamp.position = NULL, 
                           legend.position = legend.position, ...),
            fig.DHI(x, timestamp.position = NULL, 
                       legend.position = legend.position, ...), 
            # fig.azimuth(x, legend.position = legend.position, 
            #             timestamp.position = NULL, ...),
            # # fig.dPS(x),
            labels = c("GHI", "ext_irrad", "clearness_index", 
                       "diffuse_fraction", "DNI", "DHI"),
            ncol = 2, nrow = 3, hjust = -0.1, vjust = 1.1, label_size = 12,
            label_colour = "black")
}
fig.ghi_decomposition(x)

gif_merra(x, FUN = "fig.ghi_decomposition", dirname = "images", fps = 5,
          gif.width = 1.25 * 576, 
          gif.height = 1.25 * 360 * 1.21, 
          filename.prefix = "fig.ghi_decomposition_")

```
</details>


```{r fig_DHI, fig.cap="Global Horizontal Irradiance decomposition", echo=FALSE}
gif_file <- "../images/fig.ghi_decomposition_720x544_5fps.gif"
if (file.exists(gif_file)) include_graphics(gif_file)

```


#### PV-array position

<details><summary>Figure code</summary>
```{r, eval = rerun}
x <- pv_array_position(x, array.type = "all")
x

fig.array.tilt <- function(x, array.type = "fl", ...) {
  plot_merra(x, paste0("array.tilt.", array.type), limits = c(0, 90), 
             palette = "Spectral", direction = -1, ...)
}

fig.array.azimuth <- function(x, array.type = "fl", ...) {
  plot_merra(x,  paste0("array.azimuth.", array.type), limits = c(0, 360), 
             direction = -1, palette = "Set1", ...)
}

summary(x$array.tilt.fh); summary(x$array.azimuth.fh)
fig.array.tilt(x, "fh")
fig.array.azimuth(x, "fh")

summary(x$array.tilt.fl); summary(x$array.azimuth.fl)
fig.array.tilt(x, "fl")
fig.array.azimuth(x, "fl")

summary(x$array.tilt.th); summary(x$array.azimuth.th)
fig.array.tilt(x, "th")
fig.array.azimuth(x, "th", scale = x$na)

# summary(x$array.tilt.tl); summary(x$array.azimuth.tl)
# fig.array.tilt(x, "tl")
# fig.array.azimuth(x, "tl", scale = x$na)

summary(x$array.tilt.tv); summary(x$array.azimuth.tv)
fig.array.tilt(x, "tv")
fig.array.azimuth(x, "tv", scale = x$na)

summary(x$array.tilt.td); summary(x$array.azimuth.td)
fig.array.tilt(x, "td")
fig.array.azimuth(x, "td", scale = x$na)

fig.pv_array_position <- function(
  x, legend.position = c(0.81, 0.08), 
  timestamp.position = c(125, 85), ...) {
  plot_grid(
    fig.array.tilt(x, "fh", timestamp.position = NULL, legend.position = NULL, 
                   scale = x$na, ...), 
    fig.array.azimuth(x, "fh", timestamp.position = timestamp.position, 
                      legend.position = NULL, scale = x$na, ...),
    fig.array.tilt(x, "fl", timestamp.position = NULL, legend.position = NULL, 
                   scale = x$na, ...), 
    fig.array.azimuth(x, "fl", timestamp.position = NULL, legend.position = NULL, 
                      scale = x$na, ...),
    fig.array.tilt(x, "th", scale = x$na, 
                   timestamp.position = NULL, legend.position = NULL, ...), 
    fig.array.azimuth(x, "th",  scale = x$na, 
                   timestamp.position = NULL, legend.position = NULL, ...),
    # fig.array.tilt(x, "tl", scale = x$na, 
    #                timestamp.position = NULL, legend.position = NULL, ...), 
    # fig.array.azimuth(x, "tl",  scale = x$na, 
    #                timestamp.position = NULL, legend.position = NULL, ...),
    fig.array.tilt(x, "tv", scale = x$na, 
                   timestamp.position = NULL, legend.position = NULL, ...), 
    fig.array.azimuth(x, "tv", scale = x$na, 
                   timestamp.position = NULL, legend.position = NULL, ...),
    fig.array.tilt(x, "td", scale = x$na, 
                   timestamp.position = NULL, 
                   legend.position = legend.position, ...), 
    fig.array.azimuth(x, "td", scale = x$na, 
                   timestamp.position = NULL, 
                   legend.position = legend.position, ...),
  labels = c("fh - tilt", "fh - azimuth", 
             "fl - tilt", "fl - azimuth", 
             "th - tilt", "th - azimuth", 
             # "tl - tilt", "tl - azimuth",
             "tv - tilt", "tv - azimuth", 
             "td - tilt", "td - azimuth"),
            ncol = 2, hjust = -0.1, vjust = 1.1, label_size = 12,
            label_colour = "#fdf8db")
}
fig.pv_array_position(x)

# ggsave("images/solar_array_position_by_type.png", width = 7, height = 10.5)

gif_merra(x, FUN = "fig.pv_array_position", dirname = "images", fps = 5,
          gif.width = 1.25 * 576, 
          gif.height = round(1.25 * 360 * 1.21 * 5 / 3), 
          filename.prefix = "fig.pv_array_position_")

```
</details>


```{r fig_array_position, fig.cap="PV-array position for alternative tracking systems", echo=FALSE}
gif_file <- "../images/fig.pv_array_position_720x908_5fps.gif"
if (file.exists(gif_file)) include_graphics(gif_file)

```

#### Angle of Incidence (AOI)
<details><summary>Figure code</summary>
```{r, eval = rerun}
x <- angle_of_incidence(x, array.type = "all")

fig.AOI <- function(x, array.type = "fl", ...) {
  plot_merra(x, paste0("AOI.", array.type), limits = c(0, 90), 
             direction = 1, 
             # palette = "Blues", 
             palette = "Set1", 
             legend.name = "angle\u00b0",
             scale = 180/pi, ...)
}
summary(x$AOI.fh); summary(rad2deg(x$AOI.fh))
fig.AOI(x, "fh")

summary(rad2deg(x$AOI.fl))
fig.AOI(x, "fl")

summary(rad2deg(x$AOI.th))
fig.AOI(x, "th")

# summary(rad2deg(x$AOI.tl))
# fig.AOI(x, "tl")

summary(rad2deg(x$AOI.tv))
fig.AOI(x, "tv")

summary(rad2deg(x$AOI.td))
fig.AOI(x, "td")

fig.AOI.matrix <- function(x, 
                         legend.position = c(0.81, 0.08), 
                         timestamp.position = c(125, 85), 
                         ...) {
  plot_grid(
    fig.AOI(x, "fh", timestamp.position = NULL, 
            legend.position = NULL, ...), 
    fig.AOI(x, "fl", timestamp.position = timestamp.position, 
            legend.position = NULL, ...),
    fig.AOI(x, "th", timestamp.position = NULL, 
            legend.position = NULL, ...), 
    # fig.AOI(x, "tl", timestamp.position = NULL, 
    #         legend.position = legend.position, ...),
    fig.AOI(x, "tv", timestamp.position = NULL, 
            legend.position = NULL, ...), 
    fig.AOI(x, "td", legend.position = legend.position, 
            timestamp.position = NULL, ...),
            # fig.dPS(x),
            labels = c("AOI - fh", "AOI - fl", "AOI - th", 
                       # "AOI - tl", 
                       "AOI - tv", "AOI - td"),
            ncol = 2, nrow = 3, hjust = -0.1, vjust = 1.1, label_size = 12,
            label_colour = "#fdf8db")
}
fig.AOI.matrix(x)

# ggsave("images/AOI_by_type.png", width = 7, height = 8.5)

gif_merra(x, FUN = "fig.AOI.matrix", dirname = "images", fps = 5,
          gif.width = 1.25 * 576, 
          gif.height = round(1.25 * 360 * 1.21), 
          filename.prefix = "fig.AOI.matrix_")

```
</details>


```{r fig_AOI, fig.cap="PV-array position for alternative tracking systems", echo=FALSE}
gif_file <- "../images/fig.AOI.matrix_720x544_5fps.gif"
if (file.exists(gif_file)) include_graphics(gif_file)

```


#### Plane of Array (POA) irradiance 

```{r, eval = rerun}
x <- poa_irradiance(x, array.type = "all", keep.all = T)
fig.poa.grid(x)

# ggsave("images/poa_by_type.png", width = 7, height = 8.)

```

### Validation: comparison with NREL's PVWAtts model estimates

<details>
  <summary>code</summary>

**Fetching POA estimates from NREL's PVWatts**
```{r, eval=FALSE}
st_id <- merra2ools::pvwatts_stations_sample
# Fetching the data from PVWAtts in select locations
if (!file.exists("tmp/pvwatts_poa_db.RData")) {
  ll <- list()
  for (tp in c("fh", "fl", "th", "tv", "td")) {
    for (i in 1:nrow(st_id)) {
      x <- fetch_pvwatts(
        array.type = tp,
        lon = st_id$lon_stn[i],
        lat = st_id$lat_stn[i],
        # tilt = abs(st_id$lat[i]),
        # tilt = 0,
        azimuth = ifelse(st_id$lat_stn[i] < 0, 0, 180),
        radius = 100
      )
      cat("type = ", tp, ", i = ", i, ", lon = ", 
          st_id$lon_stn[i], ", lat = ", st_id$lat_stn[i], 
          ", poa: ", length(x$outputs$poa), "\n", sep = "")
      nm <- paste(tp, i, sep = "_")
      ll[[nm]] <- x
      if (!is.null(x$error$code)) {
        if (x$error$code == "OVER_RATE_LIMIT") break
      }
    }
    if (!is.null(x$error$code)) {
      if (x$error$code == "OVER_RATE_LIMIT") break
    }
  }
  names(ll)
  save(ll, st_id, file = "tmp/pvwatts_poa_full_list.RData") #'raw' data

# Merging hourly POA (PVWatts) ####
  yday_hour <- expand.grid(hour = 1:24, yday = 1:365) %>% as_tibble()
  for (i in 1:length(ll)) {
    n <- names(ll)[i]
    cat(n); cat(" ")
    x <- sapply(ll[[i]]$outputs, unlist)
    x <- try(
      as_tibble(x[5:15]) %>%
        mutate(
          # request = i,
          lon_stn = ll[[i]]$station_info$lon,
          lat_stn = ll[[i]]$station_info$lat,
          lonlat = paste0("(", round(lon_stn, 2), ", ", round(lat_stn, 2), ")"),
          elev = ll[[i]]$station_info$elev,
          tz = ll[[i]]$station_info$tz, 
          month = yday2month(yday_hour$yday),
          yday = as.integer(yday_hour$yday),
          hour = as.integer(yday_hour$hour) - 1,
          # lon_inp = as.numeric(ll[[i]]$inputs$lat),
          # lat_inp = as.numeric(ll[[i]]$inputs$lon),
          # module_type = as.integer(ll[[i]]$inputs$module_type),
          # losses = as.numeric(ll[[i]]$inputs$losses),
          # id = as.integer(gsub("[a-z_]", "", i)),
          array.type = substr(n, 1, 2),
          # array_type_pvwatts = ll[[i]]$inputs$array_type,
          tilt = as.numeric(ll[[i]]$inputs$tilt),
          azimuth = as.numeric(ll[[i]]$inputs$azimuth),
          .before = ac_annual), silent = T
    )
    if (i ==  1 || !exists("db") || is.null(db)) {
      db <- NULL
      if (all(class(x) != "try-error")) db <- x
    } else {
      if (all(class(x) != "try-error")) db <- bind_rows(db, x)
    }
  }
  db

  db %>% as.data.table()
  dim(db)   
  db <- right_join(select(st_id, locid, lonlat), db)
  save(db, file = "tmp/pvwatts_poa_db.RData")
} else {
  # loading pre-saved data
  (load("tmp/pvwatts_poa_db.RData"))
}    

# summarizing the data for comparison
db_m <- db %>% 
  group_by(lonlat, locid, month, hour, array.type) %>%
  summarise(poa = max(poa, na.rm = T)) %>%
  mutate(month = factor(month.abb[month], levels = month.abb, ordered = T),
         data = factor("PVWatts", levels = c("merra2ools", "PVWatts"), ordered = T)) %>%
  ungroup()
```

**POA figures**
```{r, eval=FALSE}
# Plot POA for select (random) locations ####
set.seed(0)
ii <- db_m$lonlat %in% sample(unique(db_m$lonlat), 4) &
  db_m$month %in% c("Mar", "Jul", "Oct", "Dec"); summary(ii)

ggplot(db_m[ii,]) +
  geom_line(aes(hour, poa, colour = array.type, linetype = array.type), size = 1) +
  scale_color_brewer(palette = "Set1", name = "Array type") +
  guides(colour = guide_legend(override.aes = list(linetype = 1:5))) +
  scale_linetype(guide = FALSE) +
  facet_grid(lonlat ~ month) + 
  labs(y = "POA irradiance, Watt/m2", x = "Hour, local time") +
  theme_bw()

ggplot(db_m[ii,]) +
  geom_line(aes(hour, poa, colour = month), size = 1) +
  scale_color_brewer(palette = "Set1", name = "Month") +
  facet_grid(lonlat ~ array.type) + 
  labs(y = "POA irradiance, Watt/m2", x = "Hour, local time") +
  theme_bw()
```

**Getting MERRA-2 data for the same locations**
```{r, eval=FALSE}
merra0 <- get_merra2_subset(locid = st_id$locid, 
                            from = "2009-01-01 00", 
                            to = "2011-12-31 23")
merra0
```

**Estimating POA on MERRA-2 data**
```{r, eval=FALSE}
x <- merra0 %>% fPOA(keep.all = T)
x <- mutate(x, 
            month = month(UTC),
            month = factor(month.abb[month], levels = month.abb, ordered = T)) %>%
  group_by(locid, month, hour) %>%
  summarise(
    POA.fh = max(POA.fh, na.rm = T),
    POA.fl = max(POA.fl, na.rm = T),
    POA.th = max(POA.th, na.rm = T),
    # POA.tl = max(POA.tl, na.rm = T),
    POA.tv = max(POA.tv, na.rm = T),
    POA.td = max(POA.td, na.rm = T)
    # SWGDN = max(SWGDN, na.rm = T)
  ) %>%
  ungroup() %>% #select(-lon, -lat) %>%
  left_join(select(st_id, locid, tz_stn)) %>%
  mutate(hour = hour_utc2tz(hour, tz_offset = tz_stn)) %>%
  select(-tz_stn) %>%
  mutate(data = factor("merra2ools", levels = c("merra2ools", "PVWatts"), ordered = T)) %>%
  pivot_longer(cols = starts_with("POA."), names_to = "array.type", values_to = "poa") %>%
  mutate(array.type = sub("POA.", "", array.type))
x
db_m[,-1]

if (F) { # interim check - POA figure based on MERRA-2
  jj <- x$locid == sample(x$locid, 1)
  xx <- left_join(x[jj,], select(st_id, locid, lonlat))
  ggplot(xx) +
    geom_line(aes(hour, poa, colour = array.type), size = 1) +
    # facet_grid(month~id) +
    facet_wrap(~month) +
    theme_bw() +
    labs(title = paste0("PVWatts station coord ", 
                        unique(xx$lonlat), 
                        ", MERRA-2 locid = ", unique(xx$locid)))
}

```

**Compare PVWatts and MERRA-2 estimates**
```{r, eval=FALSE}
y <- bind_rows(
  select(db_m, data, locid, month, hour, array.type, poa), 
  select(x, data, locid, month, hour, array.type, poa)) %>%
  left_join(st_id[, c("locid", "lonlat")]) %>%
  add_coord() %>% ungroup() %>%
  mutate(locid_lonlat = paste0(locid, " (", lon, ", ", lat, ")")) %>%
  as.data.table() %>%
  select(-lon, -lat)

fig.PVWatts_MERRA <- function(x, lid = NULL) {
  lids <- unique(x$locid)
  if (length(lids) > 1) {
    if (is.null(lid)) {lid <- sample(lids, 1)}
  }
  
  if (!is.null(lid)) {
    ii <- x$locid %in% lid
  } else {
    ii <- rep(TRUE, nrow(x))
  }
  ggplot(x[ii,]) +
    geom_line(aes(hour, poa, colour = data), size = 1) +
    facet_grid(month ~ array.type) +
    theme_bw() +
    scale_color_viridis_d(name = "") +
    theme(
      legend.position = "bottom",
          legend.margin = margin(0,0,0,0),
          legend.box.margin = margin(-10,0,0,0),
          plot.margin = unit(c(0,0,0,0), "mm"),
          plot.title = element_text(hjust = 0.5),
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    guides(colour = guide_legend(nrow = 1)) +
    labs(title = paste0("PVWatts station coord ", unique(x[ii,]$lonlat), 
                        "; MERRA-2 locid = ", unique(x[ii,]$locid)),
         x = "Hour", y = "POA, Watt/m2, monthly maximum for every hour")
}
fig.PVWatts_MERRA(y)

lids <- unique(y$locid) %>% sort()
fld <- "images/poa_pvwatts_merra"
if (!dir.exists(fld)) dir.create(fld, recursive = T)
for (i in lids) {
  cat(i, "\n")
  gg <- fig.PVWatts_MERRA(y, i)
  fname <- paste0("poa_pvwatts_vs_merra2_locid_", formatC(i, width = 6, flag = "0"), ".png")
  ggsave(file.path(fld, fname), gg, width = 6, height = 8, units = "in")
}
    
```
</details>

```{r, fig.height=8.5, fig.width=6, fig.cap="Comparison of POA estimates (merra2ools vs. PVWatts 6) for different types of PV arrays", echo=FALSE}
library("EBImage")
fld <- "../images/poa_pvwatts_merra"

# imgs <- paste0("../images/poa_pvwatts_merra/", sample(list.files(fld), 5))
imgs <- paste0("../images/poa_pvwatts_merra/", list.files(fld))
display(readImage(imgs), method = "browser")
```

## Refferences
...